Scientific Reasoning: \\ Assessment of Multimodal Generative LLMs

### 2024-12-10 docs/paper/main.tex
some description of pixtral issues

\begin{example}
\label{example:1}
\begin{itemize}
    \item \textbf{Reference:} Plants absorb sunlight to produce energy.
    \item \textbf{Candidate:} Sunlight is absorbed by plants to create energy.
\end{itemize}
\end{example}

\subsubsubsection{Cosine similarity with Sentence Transformers}
Cosine similarity using SentenceTransformer embeddings provides a robust method for evaluating model outputs in the ScienceQA dataset, focusing on the semantic alignment of generated explanations with references. It captures nuanced relationships and conceptual understanding, which gives more understanding of the generated outputs in comparison to ROUGE and BLEU scores and that is why it has more importance in final score computation. However, due to domain specificity it can miss some intricate cases like still considering high similarity of two sentences, even though it the reference can contain crucial information about carbon dioxide, while in the candidate sentence the answer was given in relation to oxygen reproduction.
\paragraph{METEOR}
METEOR (\cite{banerjee2005meteor}) evaluates text similarity based on precision, recall, and alignment of semantically meaningful components like synonyms and stems, making it well-suited for capturing the nuanced language used in scientific reasoning.
By incorporating a weighted F-score and penalties for excessive mismatches, METEOR provides a balanced evaluation of fluency and accuracy, crucial for assessing LLM-generated answers on the Science QA Dataset. Due to these balanced text evaluation, this metric has \textbf{10\%} impact on the overall score.
% \subsubsection{Cosine similarity with SciBERT}
% We extended the evaluation of cosine similarity using SciBERT (\cite{beltagy2019scibert}) embeddings. These vectors are more consistent with the domain-specific terminology and reasoning patterns present in the Science QA Dataset because it has been properly pre-trained on scientific literature.
% SciBERT generates more accurate similarity embeddings because it better reflects the complex relationships between scientific concepts than general-purpose models like those in Sentence Transformers.
% When assessing the logical and factual coherence of scientific reasoning, the structural and contextual semantics specific to scientific texts are reflected in the embeddings produced by SciBERT.
% Assessments for jobs requiring sophisticated scientific understanding are more reliable when SciBERT is used for cosine similarity since it guarantees that the evaluation metric is domain-specific.

@@

\paragraph{Pre-experiments} We were not successful with processing with small non-chat models. For the given text + image, the models would return short non-meaningful answers, like "Yes" to a "X or Y" question, or "no idea" to a completely defined task with all relevant context included. These were multimodal models below 1B on Huggingface, not tuned to be used in a chat scenario (and prepending a "You are a helpful assistant..." to the beginning of the prompt was not enough to fix this). 
We are not sure how to use those models for our purposes as of now.

@@

\paragraph{LLaVa} The exact model used is: \texttt{llava-hf/llava-1.5-7b-hf} (and there is a larger sister 13b model). It is based on LLaMA and (?) Vicuna fine-tuned on GPT-generated multimodal instruction-following data.
It expects inputs as conversation, allowing the preprocessor to convert textual input into format "USER: {request} ASSISTANT: " where the assistant has to fill in the rest. 
We encountered the following issues while applying this model version on ScienceQA data: 
\begin{itemize}
	\item although prompted like larger models, this model returned answer strings and not choice index; this may additionally raise an issue of the returned answer not being entirely part of the choice list;
	\item although prompted to generate JSON, the model would sometimes generate a different string, e.g. simply repeating part of the task. Unexpected format happened in ?? \% of all answers. We disregarded these answers, considering them all wrong, which results in a bit more pessimistic estimation of the model's performance. The metrics on the correctly formatted answers alone are also given. 
	\item the model currently does not accept inputs without an image (text-only); according to a huggingface discussion\footnote{https://huggingface.co/llava-hf/llava-1.5-7b-hf/discussions/38}, the authors claim to have accidentally removed this possibility in a recent update. We worked around this by producing placeholder empty 10x10 white images and passing them as part of the input when the image in the dataset was missing. We also appended the suffix "Ignore image" to the end of the prompt, because the model would otherwise try to hallucinate a relation between the text and the image (e.g. "Why is the sky blue" + placeholder image results in "The sky is blue on the picture because ...");
	\item the model does not really seem to support batch processing; although officially the technical possibility exists, it has been reported online \footnote{https://huggingface.co/llava-hf/llava-1.5-13b-hf/discussions/10} that the model would only take into account the first image in the batch, which makes the processing useless. We did not risk to attempt the supported or custom batching and processed the filtered validation dataset 1 datapoint by 1, which took 10+ hours on GPU for inference. 
\end{itemize}

@@

Note: I later (after running the model) understood that the original model is this one: `liuhaotian/llava-v1.5-7b`. The results may thus be different from what would be expected from the model's paper.

>>>>

%\begin{example}
\label{example:1}
\begin{itemize}
    \item \textbf{Reference:} Plants absorb sunlight to produce energy.
    \item \textbf{Candidate:} Sunlight is absorbed by plants to create energy.
\end{itemize}
%\end{example}

\paragraph{Cosine similarity with Sentence Transformers}
Cosine similarity using SentenceTransformer embeddings provides a robust method for evaluating model outputs in the ScienceQA dataset, focusing on the semantic alignment of generated explanations with references. It captures nuanced relationships and conceptual understanding, which gives more understanding of the generated outputs in comparison to ROUGE and BLEU scores and that is why it has more importance in final score computation. However, due to domain specificity it can miss some intricate cases like still considering high similarity of two sentences, even though it the reference can contain crucial information about carbon dioxide, while in the candidate sentence the answer was given in relation to oxygen reproduction.
\paragraph{METEOR}
METEOR (\cite{banerjee2005meteor}) evaluates text similarity based on precision, recall, and alignment of semantically meaningful components like synonyms and stems, making it well-suited for capturing the nuanced language used in scientific reasoning.
By incorporating a weighted F-score and penalties for excessive mismatches, METEOR provides a balanced evaluation of fluency and accuracy, crucial for assessing LLM-generated answers on the Science QA Dataset. Due to these balanced text evaluation, this metric has \textbf{10\%} impact on the overall score.
% \subsubsection{Cosine similarity with SciBERT}
% We extended the evaluation of cosine similarity using SciBERT (\cite{beltagy2019scibert}) embeddings. These vectors are more consistent with the domain-specific terminology and reasoning patterns present in the Science QA Dataset because it has been properly pre-trained on scientific literature.
% SciBERT generates more accurate similarity embeddings because it better reflects the complex relationships between scientific concepts than general-purpose models like those in Sentence Transformers.
% When assessing the logical and factual coherence of scientific reasoning, the structural and contextual semantics specific to scientific texts are reflected in the embeddings produced by SciBERT.
% Assessments for jobs requiring sophisticated scientific understanding are more reliable when SciBERT is used for cosine similarity since it guarantees that the evaluation metric is domain-specific.

@@

\paragraph{Pre-experiments} We were not successful in processing the dataset with small non-chat models. Those  models would return short meaningless answers, like "Yes" to a "X or Y" question, or "no idea" to a completely defined task with all relevant context included. These were multimodal models below 1B on Huggingface, not tuned to be used in a chat scenario (and prepending a "You are a helpful assistant..." to the beginning of the prompt was not enough to fix this).  We are not sure how to use those models for our purposes as of now.

@@

\paragraph{LLaVa 1.5 7B} The exact model used is: \texttt{llava-hf/llava-1.5-7b-hf} (and there is a larger sister 13b model). It is based on LLaMA and Vicuna, and is fine-tuned on GPT-generated multimodal instruction-following data. It expects inputs as conversation, allowing the preprocessor to convert textual input into format "USER: {request} ASSISTANT: " where the assistant has to fill in the rest.

We encountered the following issues while applying this model version on ScienceQA data (although, as we discovered later, the model had actually been trained on ScienceQA): 
\begin{itemize}
	\item although prompted like larger models, this model returned answer strings and not choice index; this may additionally raise an issue of the returned answer not being entirely part of the choice list;
	\item although prompted to generate JSON, the model would sometimes generate a different string, e.g. simply repeating part of the task. Unexpected format happened in ?? \% of all answers. We disregarded these answers, considering them all wrong, which results in a bit more pessimistic estimation of the model's performance. The metrics on the correctly formatted answers alone are also given. 
	\item the model currently does not accept inputs without an image (text-only); according to a huggingface discussion\footnote{https://huggingface.co/llava-hf/llava-1.5-7b-hf/discussions/38}, the authors claim to have accidentally removed this possibility in a recent update. We worked around this by producing placeholder empty 10x10 white images and passing them as part of the input when the image in the dataset was missing. We also appended the suffix \textit{"Ignore image"} to the end of the prompt, because the model would otherwise try to hallucinate a relation between the text and the image (e.g. "Why is the sky blue" + placeholder image results in "The sky is blue on the picture because ...");
	\item the model does not really seem to support batch processing; although officially the technical possibility exists, it has been reported online \footnote{https://huggingface.co/llava-hf/llava-1.5-13b-hf/discussions/10} that the model would only take into account the first image in the batch, which makes the processing useless. We did not risk to attempt the supported or custom batching and processed the filtered validation dataset 1 datapoint by 1. 
\end{itemize}

@@

%Note: I later (after running the model) understood that the original model is this one: `liuhaotian/llava-v1.5-7b`. The results may thus be different from what would be expected from the model's paper.
